### 2021-07-31 paper/localpsf.tex
boundary extension written okish

\section{Constructing the product-convolution H-matrix approximation}
\label{sec:method}
This section describes how we compute the convolution kernels, $\varphi_i$, and weighting functions, $w_i$, for the product-convolution approximation, $\AopPc \approx \Aop$, given in \eqref{eq:product_convolution}. We also describe how we evaluate kernel entries of the approximation, $\Phi_\text{pc}(y,x)$, how we convert $\AopPc$ to H-matrix format, and (optionally) how we modify $\AopPc$ to force it to be symmetric positive semi-definite.

@@

\subsection{Estimating impulse response supports}
\label{eq:mean_and_covariance_estimation}

@@

\subsection{Greedy sample point selection}
\label{sec:greedy_point_selection}

@@

\subsection{Computation and storage of convolution kernels}
\label{sec:get_impulse_response}

@@

To convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), we will need to evaluate $\varphi_{x_i}(z)$ for a large number of points $z$. Per \eqref{eq:varphi_eval}, evaluating $\varphi_{x_i}(z)$ requires evaluating $\eta_b(z+x_i)$. This is not ideal. In computations, $\eta^b$ will be a finite element function formed on an irregular mesh, and $z+x_i$ will typically not be a gridpoint of the mesh. Evaluating $\eta_b(z+x_i)$ therefore requires finding which mesh cell that $z+x_i$ is in, then evaluating several polynomial functions to evaluate the finite element function on that cell. Although this procedure is asymptotically scalable, it is slow in practice \cite{FEMEVALSLOW}. To speed up evaluation of $\varphi_{x_i}$, we interpolate $\varphi_{x_i}$ onto a regular rectilinear grid on a coordinate axis aligned box that just barely contains the ellipsoid $E_{x_i}$. We transfer the function $\varphi_{x_i}$ from the irregular mesh to the regular grid by evaluating $\varphi_{x_i}$ at all gridpoints in the regular grid, using the slow process just described.  This is an upfront cost which is done ``offline.'' After $\varphi_{x_i}$ has been transferred to the regular grid, we use fast regular grid interpolation to perform all subsequent ``online'' evaluations of $\varphi_{x_i}$. This yields a substantial overall speedup because the number of gridpoints in the regular grid for $\varphi_{x_i}$ is typically much smaller than the number of evaluations of $\varphi_{x_i}$ required to convert $\AopPc$ to H-matrix format. Having $\varphi_{x_i}$ on a regular grid also simplifies the implementation of the boundary impulse response extension procedure that we describe in Section \ref{sec:boundary_extension}.

@@

@@

\subsection{Addressing convolution kernel boundary artifacts}
\label{sec:boundary_extension}

Evaluating the kernel for our product-convolution approximation requires evaluating $\varphi_{x_i}(y-x)$. But this may be undefined, because $y-x$ may reside outside the support of $\varphi_{x_i}$, even if $x \in \Omega$ and $y \in \Omega$. The default approach to deal with this problem is to extend $\varphi_{x_i}$ by zero wherever it is undefined. But this leads to substantial errors, known as boundary artifacts, when the sample point $x_i$ is near the boundary of the domain. In this case, the impulse response is artificially chopped off by the domain boundary. See Figure FIG for an illustration, and see \cite{ADAPTIVECONVMYPAPER} for a more detailed discussion of boundary artifacts in.

To address boundary artifacts, we fill in missing/undefined portions of convolution kernels using values from convolution kernels for nearby sample points. This is similar to the approach used in \cite{ADAPTIVECONVMYPAPER}, except

\subsection{Poisson interpolation weighting functions}
\label{sec:weighting_functions}

@@

\subsection{Hierarchical matrix construction}
\label{sec:H_matrix_conversion}

@@

\subsection{Rational positive semi-definite modification}
\label{sec:make_spd}

>>>>

\section{Constructing the product-convolution H-matrix approximation}
\label{sec:method}
This section describes how we compute the convolution kernels, $\varphi_i$, and weighting functions, $w_i$, for the product-convolution approximation, $\AopPc \approx \Aop$, given in \eqref{eq:product_convolution}. We also describe how we convert $\AopPc$ to H-matrix format, and (optionally) how we modify $\AopPc$ to force it to be symmetric positive semi-definite.

@@

\subsection{Sample point selection}

\subsubsection{Estimating impulse response supports}
\label{eq:mean_and_covariance_estimation}

@@

\subsubsection{Greedy ellipsoid packing algorithm}
\label{sec:greedy_point_selection}

@@

\subsection{Convolution kernels}
\label{sec:get_impulse_response}

@@

\subsubsection{Interpolation onto local regular grids}

To convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), we will need to evaluate $\varphi_{x_i}(z)$ for a large number of points $z$. Per \eqref{eq:varphi_eval}, evaluating $\varphi_{x_i}(z)$ requires evaluating $\eta_b(z+x_i)$. This is not ideal. In computations, $\eta^b$ will be a finite element function formed on an irregular mesh, and $z+x_i$ will typically not be a gridpoint of the mesh. Evaluating $\eta_b(z+x_i)$ therefore requires finding which mesh cell that $z+x_i$ is in, then evaluating several polynomial functions to evaluate the finite element function on that cell. Although this procedure is asymptotically scalable, it is slow in practice \cite{FEMEVALSLOW}.

To speed up evaluation of $\varphi_{x_i}$, we interpolate $\varphi_{x_i}$ onto a regular rectilinear grid on a coordinate axis aligned box that just barely contains the ellipsoid $E_{x_i}$. We transfer the function $\varphi_{x_i}$ from the irregular mesh to the regular grid by evaluating $\varphi_{x_i}$ at all gridpoints in the regular grid, using the slow process described in the previous paragraph. This is an upfront cost which is done ``offline.'' After $\varphi_{x_i}$ has been transferred to the regular grid, we use fast regular grid interpolation to perform all subsequent ``online'' evaluations of $\varphi_{x_i}$. This yields a substantial overall speedup because the number of gridpoints in the regular grid for $\varphi_{x_i}$ is typically much smaller than the number of evaluations of $\varphi_{x_i}$ required to convert $\AopPc$ to H-matrix format. 
%Having $\varphi_{x_i}$ stored on a regular grid also simplifies the implementation of the method that we use to deal with boundary artifacts, which will be described in Section \ref{sec:boundary_extension}.

@@

When $x_i$ is near a boundary, it may occur that some gridpoints for the regular grid are in locations where $\varphi_{x_i}$ is undefined. In this case, we store a placeholder value for $\varphi_{x_i}$ at those gridpoints. We use the IEEE standard value of ``NaN'' (not a number) for this placeholder. These NaN values are later replaced with more appropriate values, as we will discuss in Section \ref{sec:boundary_extension}.

@@

\subsubsection{Addressing boundary artifacts}
\label{sec:boundary_extension}

Evaluating the kernel for our product-convolution approximation, $\AkerPc(y,x)$, requires evaluating $\varphi_{x_i}(y-x)$. But $\varphi_{x_i}(y-x)$ is undefined  when $y - x + x_i$ is outside $\Omega$, which may occur even if $x \in \Omega$ and $y \in \Omega$. The default approach to deal with this problem is to extend $\varphi_{x_i}$ by zero wherever it is undefined. But this leads to substantial errors, known as boundary artifacts, which are especially significant when the sample point $x_i$ is near the boundary of the domain. In this case, the impulse response is artificially chopped off by the domain boundary. See Figure FIG for an illustration, and see \cite{ADAPTIVECONVMYPAPER} for more detailed discussion of boundary artifacts.

While $\varphi_{x_i}$ may be undefined at a point $z$, often there is a sample point $x_j$ that is near $x_i$, but further away from the boundary than $x_i$, such that $\varphi_{x_j}$ is well defined at $z$. To address boundary artifacts, we fill in missing/undefined portions of convolution kernels using information from other convolution kernels associated with the $k$ nearest sample points. Kernels associated with nearer sample points are given precedence over kernels associated with farther away sample points. After the convolution kernel has been extended using information from neighboring convolution kernels, it is then further extended by zero as needed.

Let $x_{n_1}, x_{n_2}, \dots, x_{n_k}$ be the $k$ nearest sample points to $x_i$, ordered by nearness, including $x_i$ itself. That is, $x_{n_1}=x_i$ (the point itself), $x_{n_2}$ is the nearest sample point to $x_i$ other than $x_i$ itself, $x_{n_3}$ is the second nearest sample point to $x_i$ other than $x_i$ itself, and so on. Also, recall that in computations that the functions $\varphi_{x_i}$ are stored on regular rectilinear grids. To extend $\varphi_{x_i}$, first we form an expanded regular rectilinear grid that has the same grid spacing as the grid used for $\varphi_{x_i}$, but is large enough to contain the boxes associated with all of the neighbor kernels $\varphi_{x_{n_1}}, \varphi_{x_{n_2}}, \dots, \varphi_{x_{n_k}}$. We initialize an extended convolution kernel, $\varphi_{x_i}^E$, to zero on this expanded grid. Then we interpolate the functions $\varphi_{x_{n_j}}$ onto this expanded grid one at a time in reverse order. That is, first we interpolate the kernel, $\varphi_{x_{n_k}}$, associated with the furthest point from $x_i$ onto the grid, then the kernel, $\varphi_{x_{n_{k-1}}}$, associated with the second furthest point, and so on until we reach the original kernel $\varphi_{x_{n_1}} = \varphi_{x_i}$. At each step, we overwrite existing values of $\varphi_{x_i}^E$ with new values from the kernel currently being interpolated, except when the new value of $\varphi_{x_{n_j}}$ is undefined (``NaN''). At gridpoints where the kernel being interpolated is undefined, we leave $\varphi_{x_i}^E$ unchanged. In this way, we fill in undefined values of $\varphi_{x_i}$ with values from successively further and further away kernels. This process is illustrated in Figure FIG. To avoid notational clutter, and since we will henceforth only work with the extended kernel, $\varphi_{x_i}^E$, in the remainder of this paper the symbol $\varphi_{x_i}$ refers to $\varphi_{x_i}^E$.

\subsection{Poisson weighting functions}
\label{sec:weighting_functions}

@@

\subsection{Postprocessing the product-convolution approximation}

\subsubsection{Hierarchical matrix construction}
\label{sec:H_matrix_conversion}

@@

\subsubsection{Rational positive semi-definite modification}
\label{sec:make_spd}
